---
title: "Take-Home Midterm"
author: "Your Name Here!"
date: "10/7/2019"
output:
  pdf_document:
  html_document:
    toc: yes
    toc_float: yes
header-includes: \usepackage{xcolor}
---

# Question 1  

In 2012 data were obtained on 10,933 Okcupid users' profiles. Okcupid is a free 
dating website, designed by mathematicians, which today has over 112,000 users,
boasts that over 7.3 million messages per day, and prides themselves on the 
inclusive nature of their website (all users are welcome).  

The profiles.csv dataset provides an assortment of self reported (user input) 
variables:    

* age
* body type
* education
* ethnicity
* height (in inches)
* income
* job type
* sexual orientation
* religion
* sexual identity
* zodiac sign
* marital status

Data analytics at Okcupid uses these data to make the best possible "matches". 
The data scientists are interested if men's self reported heights differ by 
sexual orientation.  


## Part One 

Use the following code to read in the data. 

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
## Change the code below to specify the correct path to load in the data 
profiles <- read_csv("data/profiles.csv")

```

__Provide a description of the process being executed in the code below.__ 
__(e.g. What procedure is occurring on each line of code?)__ 
<!-- Your response here! Do not delete this line! -->

</br>

\color{blue}
The first line creates a new variable (object) named `male_profiles`, using the 
`profiles` dataset. The second line subsets the `profiles` dataset to only be 
profiles where the person identified as a male who's sexual orientation was not 
bisexual, and who's heights were between 48 and 84 inches (4 and 7 feet). The 
last line extracts only the `height` and `orientation` columns from the dataset. 
\color{black}

</br>
</br>

```{r}
male_profiles <- profiles %>%  
                  filter(sex == "m", orientation != "bisexual", 
                         height > 48, height < 84) %>% 
                  select(height, orientation) 
```


## Part Two 

Using the male_profiles dataset, create visualizations of the distribution of 
heights for men who identify as straight and those who identify as gay. Make 
sure that your plot has:  

* nice axis labels
* a nice title 

```{r}
## Your code to generate a plot of the distribution of heights for straight and 
## gay men goes here!  

male_profiles %>% 
  ggplot(aes(x = orientation, y = height)) + 
  geom_violin(aes(color = orientation)) + 
  theme(legend.position = "none") + 
  labs(x = "Self-Identified Sexual Orientation", 
       y = "Self-Reported Height (inches)") + 
  ggtitle("Differences in Self-Reported Height between Straight and Gay Males")

```

Interpret what you see in the plot you created. Be sure you reference the 
correct order of subtraction used!  
<!-- Your response here! Do not delete this line! -->

</br>

\color{blue}
Based on the side-by-side violin plots, it appears that the self-reported height 
of gay men has a similar center to that of straight men, but wildly different 
shapes. The distribution of self-reported height for gay men is very bell-shaped
and symmetric, with a short right tail. While the center of distribution of 
heights for straight men is similar to that of gay men, the shape of the 
distribution is very jagged, with many modes, and has very long tails. 
\color{black}

</br>
</br>

## Part Three 

The data scientists at OkCupid chose to analyze these data using a permutation 
test for a difference in mean self-reported height.   

__Do you agree with their decision? Why or why not?__  
__Be sure to reference the plot you created in Part Two!__ 
<!-- Your response here! Do not delete this line! -->

</br>

\color{blue}
Choosing a permutation test over a t-test indicates that the data scientists 
believed the assumption of normally distributed responses was violated. While 
the overall shape of the self-reported height of straight men looks normal-esk, 
I would agree that the multi-mode distribution violates this assumption. However, 
based on the long tails see in the heights of straight men, the assumption of 
equal variance may also be violated, causing an issue for using a permutation 
test! 
\color{black}

</br>
</br>


## Part Four  

Regardless of your opinion of their choice, use the male_profiles data to 
reproduce the permutation test performed by the original researchers and write 
a statistical report discussing whether men who identify as straight self-report
different heights than men who identify as gay.  You may wish to use your 
answers to previous parts and/or additional analyses as elements of your report.


```{r, cache = TRUE, echo = FALSE}
## Your code to carry out a permutation test for a difference in means goes here!  
library(infer)

diff_mean <- male_profiles %>% 
  specify(height ~ orientation) %>% 
  calculate(stat = "diff in means", 
            order = c("straight", "gay")
            )

permutations <- male_profiles %>% 
  specify(height ~ orientation) %>% 
  hypothesize(null = "independence") %>% 
  generate(type = "permute", reps = 5000) %>% 
  calculate(stat = "diff in means", 
            order = c("straight", "gay")
  )
  
get_p_value(permutations, diff_mean, direction = "two_sided")

permutations %>% 
  visualise(method = "simulation") + 
  shade_p_value(diff_mean, 
                direction = "two_sided", 
                color = "red", 
                fill = "pink")

```

## Part Five  

Based on the results of your permutation test, create a *Results* and 
*Conclusions* section for a statistical report for these data and research 
question.  

</br>

__Results__   

<!-- Your response here! Do not delete this line! -->
\color{blue}
The permutation test for a difference in mean self-reported height resulted in 
very strong evidence against the null hypothesis (p-value = 0.0108 on 5000 
simulations) of no difference in mean height between males who identified as gay
and those who identified as straight. Thus, we conclude that there is a 
difference in mean self-reported height between straight male users and gay male
users.  
\color{black}

</br>

__Conclusion__   

<!-- Your response here! Do not delete this line! -->
\color{blue}
The difference in mean self-reported height between straight male users and gay 
male users cannot be inferred to the entire population of OkCupid users, as these
data were not randomly sampled from all possible user profiles. Furthermore, we 
cannot conclude that sexual orientation causes the difference in self-reported 
height, as sexual orientation cannot be randomly assigned. Therefor, we can 
conclude that there is an association between sexual orientation and a 
difference in self-reported height for these OkCupid users. 
\color{black}

</br>
</br>

\pagebreak

# Question 2 (Paired Data)

Data on the number of males and females who entered in the New York City marathon 
have been recorded every year since 1978. These data are provided in the 
ny-marathon-2016.txt file. Social scientists are interested in if over this time
period, a larger number of females register for the NYC marathon than males. 

## Part One 

Use the following code to read in the data. 

```{r, message = FALSE}
marathon <- read_tsv("data/ny-marathon-2016.txt")

```

__What is the mutate() function doing in the code below.__ 
__(e.g. How is it modifying the marathon dataset?)__ 


```{r}
marathon <- marathon %>% 
            rename(Male = Men, Female = Women) %>% 
            mutate(Diff = Female - Male)
```

<!-- Your response here! Do not delete this line! -->
</br>

\color{blue}
The `mutate()` function is creating a new variable, named `Diff`, which is the
difference between the `Male` and `Female` columns (`Female` - `Male`). 
\color{black}

</br>
</br>


## Part Two 

Create a visualization of how the distribution of differences in the number of 
male and female runners over the last 40 years. Make sure that your plot has:  

* nice axis labels
* a nice title

```{r, warning = FALSE}
## Your code to generate a plot of the distribution of differences in male and 
## female NYC runners goes here!  

marathon %>% 
  ggplot(aes(x = Diff)) + 
  geom_density(fill = "tomato") + 
  labs(x = "Difference in NYC Marathon Runners (Female - Male)", 
       y = "Density") + 
  ggtitle("Differences in Number of NYC Marathon Runners (Female - Male), 
          for years 1978 to 2016")

```

Interpret what you see in the plot you created. Be sure you reference the correct
order of subtraction used!  
<!-- Your response here! Do not delete this line! -->

</br>

\color{blue}
Based on the density plot, it appears that most frequently there are around 17 
more female runners who register for the NYC marathon than male runners. The 
maximum difference between registrants was about 24 and the minimum difference
was nearly 11 (more females than males). Overall, the density plot appears 
fairly bell-shaped and symmetric, with a short, left tail and a long, right 
tail.
\color{black}

</br>
</br>


## Part Three 

Using the plot you created in Part Two and the Q-Q plot below, assess whether 
the assumptions of conducting a t-test are violated. Be sure to address **every**
assumption associated with conducting a t-test. Justify your decisions based on 
the information provided about these data and the visualizations. 

```{r}
run_lm <- lm(Diff ~ 1, data = marathon)
plot(run_lm, which = 2)

```
<!-- Your response here! Do not delete this line! -->

</br>

\color{blue}
Based on the density plot above and the Normal Q-Q plot, I do not believe the 
assumption of normally distributed responses is violated. As we saw previously, 
the distribution of differences appears fairly bell-shaped and symmetric and 
the Q-Q plot show that only 1 observation deviates substantially from the line. 

The assumption of independence no longer causes concern, after taking the 
difference of the runners, as we would expect the number of female runners one 
year to provide information as to how many we could expect the next year. 
\color{black}

</br>
</br>


## Part Four  

Using the marathon dataset, carry out a t-test to investigate if, over this time
period, a larger number of males register for the NYC marathon than women. 

```{r}
## Your code to carry out a t-test for a single mean goes here!  
library(infer)

marathon %>% 
  t_test(Diff ~ NULL, 
         alternative = "greater", 
         mu = 0)

```


## Part Five 

Based on the results of your t-test test, create a *Results* and 
*Conclusions* section for a statistical report for these data and research 
question.  

</br>

__Results__ 
<!-- Your response here! Do not delete this line! -->

\color{blue}
The t-test resulted in very strong evidence against the null hypothesis
(t-stat = 42.7 on 37 degrees of freedom, p-value $<$ 0.0001) of no difference in 
number of male and female runners who register for the NYC marathon. Therefore, 
we conclude that a greater number of female runners register for the NYC 
marathon than male runners. 
\color{black}

</br>

__Conclusion__ 
<!-- Your response here! Do not delete this line! -->

\color{blue}
The findings of these data cannot be generalized to other large scale marathons, 
as this race was not randomly chosen from all large marathons. However, we can 
generalize these findings to the years 1978-2016 of the NYC marathon, as we have
all of the data for these years (a census). We also cannot conclude that being 
a female causes you to register for the NYC marathon more often, but can say 
there is an association between being a female and being more predisposed to 
registering for the NYC marathon. 
\color{black}





