---
title: "Take-Home Midterm"
author: "Your Name Here!"
date: "10/7/2019"
output: pdf_document
---

General instructions:  

* Be sure that figures and tables are clearly labeled and captioned -- this may 
mean not using R defaults!
* When in doubt, answer with a sentence; streams of code and output **do not** 
constitute a complete answer.
* And finally, all work should be your own -- no discussion or collaboration 
with classmates.

Submission instructions:

We will be using Gradescope to collaboratively grade midterm exams.  You should have received an email when the class was created on Gradescope; contact Dr. Carnegie or Dr. Green **no later than 2 days before exam is due** if you need the message resent.

Gradescope requires a pdf file for submission.  We are hosting the exam files on the RStudio Cloud so you can all knit to pdf there, then download the knitted file and upload to Gradescope.

Instructions for turning this assignment in to Gradescope:  

1)	Log-in to Gradescope.
2)	Enter the course Stat 411/511
3)	Follow the instructions provided here to submit this to Midterm Take-home:
https://www.gradescope.com/help#help-center-item-student-submitting

Note you will need to identify which page of your pdf contains the answer to each question; this may take a couple of minutes, so be sure to allow time before the deadline to complete the submission process.

# Question 1  

In 2012 data were obtained on 10,933 Okcupid users' profiles. Okcupid is a free 
dating website, designed by mathematicians, which today has over 112,000 users,
boasts that over 7.3 million messages per day, and prides themselves on the 
inclusive nature of their website (all users are welcome).  

The profiles.csv dataset provides an assortment of self reported (user input) 
variables:    

* age
* body type
* education
* ethnicity
* height (in inches)
* income
* job type
* sexual orientation
* religion
* sexual identity
* zodiac sign
* marital status

Data analytics at Okcupid uses these data to make the best possible "matches" between users. 
Their data scientists are interested to know if men's self reported heights differ by 
sexual orientation.  


## Part One 

Use the following code to read in the data. 

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
## Change the code below to specify the correct path to load in the data 
profiles <- read_csv("data/profiles.csv")

```

__Provide a description of the process being executed in the code below.__ 
__(e.g. What procedure is occurring on each line of code?)__ 
<!-- Your response here! Do not delete this line! -->




```{r}
male_profiles <- profiles %>% 
                  filter(sex == "m", orientation != "bisexual", 
                         height > 48, height < 84) %>% 
                  select(height, orientation) 
```


## Part Two 

Using the male_profiles dataset, create visualizations of the distribution of 
heights for men who identify as straight and those who identify as gay. Make 
sure that your plot has:  

* nice axis labels
* a nice title 

```{r}
## Your code to generate a plot of the distribution of heights for straight and 
## gay men goes here!  

```

__Interpret what you see in the plot you created.__ 
<!-- Your response here! Do not delete this line! -->




## Part Three 

The data scientists at OkCupid chose to analyze these data using a 
permutation test for a difference in mean self-reported height.   

__Do you agree with their decision? Why or why not?__  
__Be sure to reference the plot you created in Part Two!__ 
<!-- Your response here! Do not delete this line! -->




## Part Four  

The following is the report generated by these data scientists. 

__Thoroughly read the report and provide a critique of their analysis and conclusions.__

*The permutation test resulted in significant evidence of an association between male OkCupid users' sexual orientation and a differnce in their self-reported height (p-value < 0.01). Therefore, we conclude that height should be included in the matchmaking algorithm differently when creating matches for straight and gay male OkCupid users.*   




```{r, cache = TRUE, echo = FALSE}
## Your code to carry out a permutation test for a difference in means goes here!  
library(infer)
set.seed(1234)

diff_mean <- male_profiles %>% 
  specify(height ~ orientation) %>% 
  calculate(stat = "diff in means", 
            order = c("straight", "gay")
            )

permutations <- male_profiles %>% 
  specify(height ~ orientation) %>% 
  hypothesize(null = "independence") %>% 
  generate(type = "permute", reps = 5000) %>% 
  calculate(stat = "diff in means", 
            order = c("straight", "gay")
  )
  
p_value <- get_p_value(permutations, diff_mean, direction = "two_sided")

permutations %>% 
  visualise(method = "simulation") + 
  shade_p_value(diff_mean, 
                direction = "two_sided", 
                color = "red", 
                fill = "pink") + 
  labs(x = "Difference in Mean Height (Gay-Straight)") + 
  ggtitle("Permutation Distribution for a Difference in Mean Self-Reported \nHeight in Male OkCupid Users")

```

<!-- Your critique of the report goes here! Do not delete this line! -->
<!-- Remember, a critique includes BOTH positive and negative attributes. -->
__Critique__:




# Question 2 

Data on the number of males and females who entered the New York City marathon 
have been recorded every year since 1978. These data are provided in the 
ny-marathon-2016.txt file. Social scientists wish to know whether, over this time
period, a larger number of males register for the NYC marathon than females. 

## Part One 

Use the following code to read in the data.

```{r, message = FALSE}
marathon <- read_tsv("data/ny-marathon-2016.txt")

```

__What is the mutate() function doing in the code below.__ 
__(e.g. How is it modifying the marathon dataset?)__ 


```{r}
marathon <- marathon %>% 
            rename(Male = Men, Female = Women) %>% 
            mutate(PctFem = Female/(Male + Female))
```

<!-- Your response here! Do not delete this line! -->




## Part Two 

Create a visualization of how the percent of runners who are female has
varied over the last 40 years. Make sure that your plot has:  

* nice axis labels
* a nice title

```{r, warning = FALSE}
## Your code to generate a plot of the percent of 
## female NYC runners goes here!  


```

__Interpret what you see in the plot you created.__   
<!-- Your response here! Do not delete this line! -->




## Part Three 

Create diagnostic plots to assess whether the assumptions required to conduct a 
t-test are violated. Be sure to address **every** assumption associated with 
conducting a t-test. Justify your decisions based on the information provided 
about these data and your visualizations. 

```{r}
#Your code to generate diagnostic plots goes here

```
<!-- Your response here! Do not delete this line! -->




## Part Four  

What type of t-test would you use to decide whether a larger
number of males register for the NYC marathon than females? Explain your reasoning.

<!-- Your response here! Do not delete this line! -->




## Part Five 

Based on the results above and any additional analyses required,
write a statistical report addressing the question of whether more 
males run the New York marathon than females.  

```{r}
#Implement additional analyses required here

```

<!-- Your statistical report goes here! Do not delete this line! -->





